# Posterior Inference & Prediction

What are the chances that the modern artist is Gen X or even younger, i.e., born in 1965 or later? let $\pi$ denote the proportion of artists represented in major U.S. modern art museums that are Gen X or younger.


```{r}
#| message: false
# Load packages
library(bayesrules)
library(tidyverse)
library(rstan)
library(bayesplot)
library(broom.mixed)
library(janitor)

# Load data
data("moma_sample")
```


```{r}
moma_sample %>% 
  count(genx)
```

So, by theory, the posterior pdf,

$$
\begin{split}
Y | \pi & \sim \text{Bin}(100, \pi) \\
\pi & \sim \text{Beta}(4, 6) \\
\end{split} \;\;\;\; \Rightarrow \;\;\;\; \pi | (Y = 14) \sim \text{Beta}(18, 92)
$$
There are three common tasks in posterior analysis: 

- Estimation, 
- Hypothesis testing, and 
- Prediction


## Posterior analysis with MCMC

### Posterior simulation

```{r}
#| results: hide
# STEP 1: DEFINE the model
art_model <- "
  data {
    int<lower = 0, upper = 100> Y;
  }
  
  parameters {
    real<lower = 0, upper = 1> pi;
  }
  
  model {
    Y ~ binomial(100, pi);
    pi ~ beta(4, 6);
  }
"

# STEP 2: SIMULATE the posterior
art_sim <- stan(model_code = art_model, data = list(Y = 14),
                chains = 4, iter = 5000*2, seed = 84735)
```


```{r}
# Parallel trace plots
mcmc_trace(art_sim, pars = "pi")
```

```{r}
#| layout-ncol: 2
# Parallel density plots
mcmc_dens_overlay(art_sim, pars = "pi")

# Autocorrelation plot
mcmc_acf(art_sim, pars = "pi")
```

```{r}
neff_ratio(art_sim, pars = "pi")
rhat(art_sim, pars = "pi")
```

Our simulation is extremely stable.


### Posterior estimation & hypothesis testing

```{r}
broom.mixed::tidy(art_sim, conf.int = TRUE, conf.level = 0.95)
```

> The estimate reports that the **median** of our 20,000 Markov chain value.

```{r}
# Shade in the middle 95% interval
mcmc_areas(art_sim, pars = "pi", prob = 0.95)
```

```{r}
# Store the 4 chains in 1 data frame
art_chains_df <- as.data.frame(art_sim, pars = "pi")
dim(art_chains_df)
```

```{r}
# Calculate posterior summaries of pi
art_chains_df %>% 
  summarise(
    post_mean = mean(pi),
    post_median = median(pi),
    post_mode = bayesrules::sample_mode(pi),
    lower_95 = quantile(pi, 0.025),
    upper_95 = quantile(pi, 0.975)
  )
```

Testing the claim that fewer than 20% of major museum artists are Gen X,

```{r}
art_chains_df %>% 
  mutate(test = pi < 0.20) %>% 
  tabyl(test)
```

By this approximation, there’s an 85% chance that Gen X artist representation is under 0.20.


### Posterior prediction    

```{r}
set.seed(1)

art_chains_df %>% 
  mutate(y_predict = rbinom(length(pi), size = 20, prob = pi)) -> art_chains_df

art_chains_df %>% 
  head()
```

```{r}
# Plot the 20,000 predictions
art_chains_df %>% 
  ggplot(aes(x = y_predict)) +
  stat_count()
```

```{r}
# Posterior mean predicton and posterior prediction interval
art_chains_df %>% 
  summarise(
    mean = mean(y_predict),
    lower_80 = quantile(y_predict, 0.1),
    upper_80 = quantile(y_predict, 0.9)
  )
```

